```{r}
library(tidycensus)
library(tidyverse)
library(keyring)


# keyring

# key_set("CENSUS_API_KEY_SECURE")
# 
# key_get("CENSUS_API_KEY_SECURE")
```

# Get Data

```{r}
age10 <-get_decennial(geography = "state",variables = "P013001",year = 2010,key =key_get("CENSUS_API_KEY_SECURE"))

```

```{r}
age10%>%ggplot(aes(x = value, y =fct_reorder(NAME, value)))+geom_point()
```

```{r}
acs18_5_vars <-load_variables(2018, "acs5", cache = TRUE)
acs18_5_vars%>%
  filter(str_detect(concept, "MEDIAN AGE BY SEX"))%>%
  head()
```

# Mapping

```{r}
library(sf) #https://r-spatial.github.io/sf/
library(tigris) #https://cran.r-project.org/web/packages/tigris/tigris.pdf
options(tigris_use_cache = TRUE)
library(viridis) #a color palette
```

```{r}

baltimore <-get_acs(state = "MD",
                    county = "Baltimore City",
                    geography = "tract",
                    variables = "B19013_001",
                    geometry = TRUE,
                    key =Sys.getenv("CENSUS_API_KEY"))

head(baltimore)

```
```{r}
baltimore %>%
  ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  coord_sf(crs = 26911) + 
  scale_fill_viridis_c(option = "magma")
```




```{r}
racevars <- c(White = "P005003",
              Black = "P005004",
              Asian = "P005006",
              Hispanic = "P004003")
baltimore_race <- get_decennial(state = "MD", county = "Baltimore City",
```


```{r}
geography = "tract",
                                variables = racevars,
                                geometry = TRUE,
                                summary_var = "P001001",
                                key =Sys.getenv("CENSUS_API_KEY"))

head(baltimore_race)
```

```{r}
baltimore_race %>% 
  mutate(pct = 100*(value/summary_value)) %>% 
  ggplot(aes(fill = pct)) +
  facet_wrap(~variable) + 
  geom_sf(color = NA) + 
  coord_sf(crs = 26915) + 
  scale_fill_viridis_c()
```

# Interactive using HTML

```{r}
library(mapview)

mapview(baltimore, zcol = "estimate")
```

# DC Example

```{r}

dc_median <- get_acs(state = "DC", county = "District of Columbia",
                     geography = "tract",
                     variables = "B19013_001",
                     geometry = TRUE,
                     key=Sys.getenv("CENSUS_API_KEY"))

dc_median %>% 
  mutate(CENSUS_TRACT = str_sub(GEOID,6,11)) ->
  dc_median

head(dc_median)

```

```{r}
dc_median %>% 
  ggplot(aes(fill = estimate) )+ 
  geom_sf(aes(geometry = geometry),color = NA) + 
  coord_sf(crs = 26915) + 
  scale_fill_viridis_c()
```

```{r}
crime <-read_csv("./data/dc-crimes-search-results.csv",
                 col_types =cols(DISTRICT =col_factor(),
                                 WARD =col_factor(),
                                 PSA =col_factor(),
                                 METHOD=col_factor()
                                 ))
```


# Joining Geometry across geometry - using csv file data for crime
```{r}
crime %>% 
  filter(METHOD!="others") %>% 
  group_by(CENSUS_TRACT, METHOD) %>% count() %>% 
  left_join(dc_median) %>%
  ggplot(aes(fill = n)) +
  geom_sf(aes(geometry = geometry),color = NA) +
  coord_sf(crs = 26915) + 
  scale_fill_viridis_c() +
  facet_wrap(~METHOD)
```

```{r}
wards <- read_sf("./data/Ward_from_2012")
wards %>% 
  mutate(WARD =parse_factor(as.character(WARD))) ->
  wards

head(wards)
```


# same geometry equals geometry joining by ward
```{r}
crime %>% 
  filter(METHOD!="others") %>% 
  group_by(WARD, METHOD) %>% 
             count() %>% 
             left_join(wards) %>% 
             ggplot(aes(fill = n)) + 
             geom_sf(aes(geometry = geometry),color = NA) + 
             coord_sf(crs = 26915) + 
             scale_fill_viridis_c() + 
             facet_wrap(~METHOD)
```

```{r}
crime %>% 
filter(METHOD!="others") %>% 
group_by(WARD, METHOD) %>% 
count() %>% 
ggplot(aes(x = WARD, y = n, color = METHOD)) + 
geom_point()
```

```{r}
au_latlong <- data.frame(longitude = -77.0888, latitude = 38.9375)
au_latlong <- st_as_sf(au_latlong, coords =c("longitude", "latitude"),
                       crs = 4326, agr = "constant")
```

# add American University to the Map

```{r}
crime %>% 
  filter(METHOD!="others") %>% 
  group_by(WARD, METHOD) %>% 
  count() %>% 
  left_join(wards) %>% 
  ggplot(aes(fill = n)) + 
    geom_sf(aes(geometry = geometry),color = NA) + 
    coord_sf(crs = 26915) + 
    scale_fill_viridis_c() + 
    facet_wrap(~METHOD) + 
    geom_sf(data = au_latlong, size = 4, shape = 23, fill = "red")
```

# GEOID more international, FIPS for US
```{r}
library(lubridate)
library(usmap)
data(statepop)
head(statepop)
```

```{r}
states <- read_csv("./data/state_geo_data.csv", col_types = cols(OID = col_character()))
```

```{r}
df_all <- read_rds("./data/covid_data.rds")

head(df_all)
```
# Using US Map package - designed for quicker maps and custom type - automatically puts Hawaii and Alaska underneath - can't zoom in or out or put long or lat

```{r}
df_all %>% 
  group_by(UID) %>% 
  filter(!is.na(cFIPS), County !="Unassigned", !str_detect(County, "^Out of ")) %>% 
  mutate(Population = max(Population),
         Daily_Total_percap = Daily_Total/Population)->
  df_all
write_csv(df_all, "./data/covid_us_data.csv")
```

```{r}
plot_usmap(regions = "states", labels = TRUE) + 
  labs(title = "US States",
       subtitle = "This is a blank map of the states of the United States.") + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))
```

```{r}
plot_usmap(include = .south_atlantic, labels = TRUE) + 
  labs(title = "US South Atlantic States",
       subtitle = "This is a blank map of the states in the South Atlantic Census Region") + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))
```

```{r}
plot_usmap(regions = "counties") + 
  labs(title = "US Counties",
       subtitle = "This is a blank map of the counties of the United States.") + 
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))
```

```{r}
plot_usmap("counties", include = c("VA", "MD", "DC")) +
  labs(title = "Custom DMV Area",
       subtitle = "These are the counties in MD, DC, and VA.")+ 
  theme(panel.background = element_rect(color = "blue", fill = "lightblue"))
```

```{r}
dmv_county_fips <- c("11001","24003","24009","24017","24021","24027","24031",
                     "24033","51013","51059","51061","51107","51153","51179",
                     "51510","51600","51610","51630","51683","51685")
plot_usmap("counties", include = dmv_county_fips, labels = TRUE) +
  labs(title = "Greater DMV Area",
       subtitle = "These are the Counties/Cities in the Greater DMV area.")+ 
  theme(panel.background = element_rect(color = "blue", fill = "lightblue"))
```

```{r}
df_all %>% 
  filter(Date == max(Date), Population >0, !is.na(sFIPS)) %>% 
  group_by(sFIPS, case_type, Date) %>% 
  summarise(Current_Total = sum(Daily_Total),
            Current_Total_percap = sum(Current_Total)/sum(Population))->
  state_totals
```

```{r}
  state_totals %>% 
    group_by(case_type) %>% 
  summarize(sumc = sum(Current_Total), n=n())
```

```{r}
state_totals %>% 
  filter(case_type =="confirmed_US") %>% 
  rename(fips = sFIPS)-> 
  state_cases

plot_usmap(data = state_cases, values = "Current_Total", color = "blue") + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Cases", label = scales::comma)+
  labs(title = "US States",
       subtitle = paste0("Total Cases by State as of ", 
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```
```{r}
state_totals %>% 
  filter(case_type =="deaths_US") %>% 
  rename(fips = sFIPS)-> 
  state_cases
plot_usmap(data = state_cases, values = "Current_Total", color = "blue") + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Deaths", label = scales::comma)+
  labs(title = "US States",
       subtitle = paste0("Total Deaths by State as of ", max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```


```{r}
state_totals %>% 
  filter(case_type =="confirmed_US") %>% 
  rename(fips = sFIPS)-> 
  state_cases

plot_usmap(data = state_cases, values = "Current_Total_percap", 
           color = "blue") + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Cases per Capita", 
                        label = scales::comma)+
  labs(title = "US States",
       subtitle = paste0("Cases per Capita by State as of ",
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "right")
```
```{r}
df_all %>% 
  filter(Date == max(Date), Population >0, !is.na(cFIPS)) %>% 
  group_by(cFIPS, case_type, Date) %>% 
  summarise(Current_Total = sum(Daily_Total),
            Current_Total_percap = sum(Daily_Total)/sum(Population))->
  county_totals
```

```{r}
county_totals %>% 
  filter(case_type =="confirmed_US") %>% 
  mutate(Current_Total_log2 = log2(Current_Total)) %>% 
  rename(fips = cFIPS)-> 
  county_cases

plot_usmap(data = county_cases, include = c("NY","NJ"),
           values = "Current_Total", color = "blue") + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Confirmed Cases", label = scales::comma)+
  labs(title = "US States",
       subtitle = paste0("Total Cases by County as of ", 
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```

```{r}
plot_usmap(data = county_cases, include = dmv_county_fips, labels = TRUE,
           values = "Current_Total", color = "blue" ) + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Cases", label = scales::comma)+
  labs(title = "DMV Region",
       subtitle = paste0("Cases by County/City as of ", 
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```
```{r}
county_totals %>% 
  filter(case_type =="deaths_US") %>% 
  rename(fips = cFIPS)-> 
  county_deaths
plot_usmap(regions = "counties", data = county_deaths, include = dmv_county_fips,
           values = "Current_Total", color = "blue",
           labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Deaths", label = scales::comma)+
  labs(title = "DMV Region",
       subtitle = paste0("Deaths by County/City as of ", 
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```
```{r}
plot_usmap(data = county_cases, include = dmv_county_fips,
           values = "Current_Total_percap", color = "blue") + 
  scale_fill_continuous(low = "white", high = "red", 
                        name = "Cases Per Capita", label = scales::comma)+
  labs(title = "DMV Region",
       subtitle = paste0("Cases per Capita by County/City as of ", 
                         max(state_cases$Date))) + 
  theme(panel.background = element_rect(color = "black", fill = "white")) +
  theme(legend.position = "top")
```
```{r}
df_all %>% 
  filter(!Daily_Total ==0) %>% 
  select(Province_State, County, cFIPS, Date, case_type, Daily_Total) %>% 
  pivot_wider(names_from = case_type, values_from = Daily_Total, values_fill = 0) %>% 
  rename("cases" = confirmed_US, "deaths" = deaths_US, "fips" = cFIPS) %>%
  arrange(Province_State, County, Date) %>% 
  group_by(Province_State, County) %>% 
  mutate(first_date = min(Date), 
         first_cases = cases[Date == first_date], 
         first_deaths = deaths[Date == first_date],
         daily_cases = c(min(first_cases),diff(cases)), 
         daily_deaths = c(min(first_deaths),diff(deaths))) %>% 
  select(Date, County, Province_State, fips, cases, daily_cases, deaths,  daily_deaths) ->
  df_daily_data
```

```{r}
dmv_county_fips <- c("11001","24003","24009","24017","24021","24027","24031",
                         "24033","51013","51059","51061","51107","51153","51179",
                         "51510","51600","51610","51630","51683","51685")
df_daily_data %>% 
  group_by(Province_State, County, fips, Date) %>% 
  summarize(max_cases = max(daily_cases)) ->
  max_cases_df
```

# Daily Change

```{r}
  plot_usmap(data = max_cases_df,include = dmv_county_fips,
               values = "max_cases", color = "blue") + 
      scale_fill_continuous(low = "white", high = "red", 
                            name = "max_cases", label = scales::comma)+
      labs(title = "DMV Region",
           subtitle = paste0("Total Deaths per Capita by County/City as of ", 
                             max(max_cases_df$Date))) + 
      theme(panel.background = element_rect(color = "black", fill = "white")) +
      theme(legend.position = "top")
```
```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(rgeos)
library(ggspatial)
library(sf)

theme_set(theme_bw())
```

```{r}
ggplot(data = world) +
  geom_sf() +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("World map", 
        subtitle = paste0("(", length(unique(world$name)), " countries)"))
```
```{r}
ggplot(data = world) +
  geom_sf(aes(fill = pop_est)) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```
```{r}
ggplot(data = world) +
    geom_sf() +
    coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

```{r}
ggplot(data = world) +
    geom_sf(aes(fill = gdp_md_est)) +
    coord_sf(xlim = c(70, 150), ylim = c(15, 55), expand = FALSE) +
    scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```


```{r}
world_points<- st_centroid(world)
world_points <- cbind(world, st_coordinates(st_centroid(world$geometry)))

ggplot(data = world) +
geom_sf() +
geom_text(data= world_points,aes(x=X, y=Y, label=name),
    color = "darkblue", fontface = "bold", check_overlap = FALSE) +
annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
    fontface = "italic", color = "grey22", size = 6) +
coord_sf(xlim = c(-97.15, -70.12), ylim = c(7.65, 30.97), expand = FALSE)
```

```{r}
ggplot(data = world) + 
  geom_sf(fill= "antiquewhite") + 
  geom_text(data= world_points,aes(x=X, y=Y, label=name), 
            color = "darkblue", fontface = "bold", check_overlap = FALSE) + 
  annotate(geom = "text", x = -90, y = 26, label = "Gulf of Mexico", 
           fontface = "italic", color = "grey22", size = 6) +
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true", 
                         pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), 
                         style = north_arrow_fancy_orienteering) + 
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  xlab("Longitude") + ylab("Latitude") + 
  ggtitle("Map of the Gulf of Mexico and the Caribbean Sea") + 
  theme(panel.grid.major = element_line(color = gray(.5), 
                                        linetype = "dashed", size = 0.5),
        panel.background = element_rect(fill = "aliceblue"))
```
```{r}
ggsave("./output/map.pdf")
ggsave("./output/map_web.png", width = 6, height = 6, dpi = "screen")
```

```{r}

(sites <- data.frame(longitude = c(-80.144005, -80.109), 
                     latitude = c(26.479005,   26.83)))

```

```{r}
sites <- st_as_sf(sites, 
                  coords = c("longitude", "latitude"),
                  crs = 4326, agr = "constant")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```
# Maps library

```{r}
library(maps)
```

```{r}
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
head(states)
```

```{r}
states <- cbind(states, st_coordinates(st_centroid(states)))
```
 
```{r}
states$ID <- stringr::str_to_title(states$ID)
```

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) + 
    geom_text(data = states, aes(X, Y, label = ID), size = 5) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

```{r}
states$nudge_y <- -1
states$nudge_y[states$ID == "Florida"] <- 1.1
states$nudge_y[states$ID == "South Carolina"] <- .1
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = states, fill = NA) + 
    geom_text(data = states, aes(X, Y, label = ID), size = 5, fontface = "bold", 
        nudge_y = states$nudge_y) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 34.3), expand = FALSE)
```

```{r}
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))
counties %>% 
  filter(str_detect(ID, "florida")) ->
  counties
counties$area <- as.numeric(st_area(counties))
head(counties)
```

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, aes(fill = area)) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```
```{r}
flcities <- tibble(state = rep("Florida", 5), 
                   city = c("Miami", "Tampa", "Orlando", "Jacksonville", "Sarasota"), 
                   lat = c(25.7616798, 27.950575, 28.5383355, 30.3321838, 27.3364347), 
                   lng = c(-80.1917902, -82.4571776, -81.3792365, -81.655651, -82.5306527))

flcities <- st_as_sf(flcities, coords = c("lng", "lat"), remove = FALSE,
                     crs = 4326, agr = "constant")

ggplot(data = world) +
  geom_sf() +
  geom_sf(data = counties, fill = NA, color = gray(.5)) +
  geom_sf(data = flcities) +
  geom_text(data = flcities, aes(x = lng, y = lat, label = city), 
    size = 3.9, col = "black", fontface = "bold") +
  coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

```{r}
library("ggrepel")
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = counties, fill = NA, color = gray(.5)) +
    geom_sf(data = flcities) +
    geom_text_repel(data = flcities, aes(x = lng, y = lat, label = city), 
        fontface = "bold", nudge_x = c(1, -1.5, 2, 2, -1), nudge_y = c(0.25, 
            -0.25, 0.5, 0.5, -0.5)) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE)
```

```{r}
ggplot(data = world) +
    geom_sf(fill = "antiquewhite1") +
    geom_sf(data = counties, aes(fill = area)) +
    geom_sf(data = states, fill = NA) + 
    geom_sf(data = sites, size = 4, shape = 23, fill = "darkred") +
    geom_sf(data = flcities) +
    geom_text_repel(data = flcities, aes(x = lng, y = lat, label = city), 
        fontface = "bold", nudge_x = c(1, -1.5, 2, 2, -1), nudge_y = c(0.25, 
            -0.25, 0.5, 0.5, -0.5)) +
    geom_label(data = states, aes(X, Y, label = ID), size = 5, fontface = "bold", 
        nudge_y = states$nudge_y) +
    scale_fill_viridis_c(trans = "sqrt", alpha = .4) +
    annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
    coord_sf(xlim = c(-88, -78), ylim = c(24.5, 33), expand = FALSE) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Observation Sites", subtitle = "(2 sites in Palm Beach County, Florida)") +
    theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"))
```

